All Samples Model

### 2018-03-16 AllSamplesModel.Rmd
...

all.patient.info = read.patient.info(patient.file, demo.file, set='all')$info
head(all.patient.info)

@@

sum.patient.data = subset(sum.patient.data, Hospital.Research.ID != 'AHM0320')

pander(sum.patient.data[,c('Hospital.Research.ID', 'Status', 'start.year', 'end.year','total.samples','total.endos','highest.path')], justify='left', caption='Training patient set')

@@

ggsave('plots/per_sample_cx.png', plot=pv, width=4, height=4, units='in', dpi=300)
ggsave('plots/per_sample_cx_roc.png', plot=roc.plot(rocX, title='Per sample cx') + plot.theme , width=4, height=4, units='in', dpi=300)

@@

```{r model, message=F, warning=F, echo=F, fig.height=16, fig.width=16}
folds = 10; splits = 5 
# file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
# if (file.exists(file)) {
#   message(paste("loading", file))
#   load(file, verbose=T)
# } else {
#   stop(paste("No model file", file, "found"))
# }

@@

```{r model_perf, message=F, warning=F, echo=F, fig.width=6, fig.height=6}
classification.rate = as_tibble(matrix(ncol=5, nrow=0, dimnames=list(c(), c('model','mean','sme','stable.coef', 'all.coef'))))

@@

ggsave('plots/performance/ridge_v_lasso.png', plot=mp$plot, width=5, height=5, units='in', dpi=300)

@@

ggsave('plots/performance/ridge_v_lasso_nohgd.png', plot=mp$plot, width=5, height=5, units='in', dpi=300)

@@

ggsave('plots/performance/no_lgd_perf.png', plot=mp$plot, height=5, width=5, units='in', dpi=300)

@@

```{r loonohgd, echo=F, message=F, warning=T, fig.width=5, fig.height=5, eval=F}
file = paste(cache.dir, 'loonohgd.Rdata', sep='/')
if (file.exists(file)) {
  message(paste("loading file", file))
  load(file, verbose=T)
} else {
  stop(paste("No file", file, "found"))
}

pg.sampNOHGD = do.call(rbind, pg.sampNOHGD)

perf = cbind.data.frame(perf=performance.at.1se, pt=unique(pg.sampNOHGD$Patient))

p = ggplot(melt(perf), aes(x=variable,y=value, group=variable, fill=variable)) +
    geom_boxplot(fill='lightcoral', color='darkgrey', outlier.fill=NA, outlier.color=NA, size=0.8) +
    geom_jitter(color='grey39', width=0.3) + plot.theme + 
    geom_label(data=melt(subset(perf, perf %in% range(perf) )), aes(x=variable, y=value, label=round(value, 3)), fontface='bold', fill='lightcoral') +
  #geom_text_repel(aes(label=pt)) +
    theme(legend.position = 'none') + labs(x='LOO model', y='Classification rate', title='Performance for LOO', subtitle='HGD samples excluded model')
p
ggsave('plots/performance/loo_perf_nohgd.png', plot=p, width=5, height=5, units='in', dpi=300)

rocNOHGD = pROC::roc(Status ~ Prediction, data=pg.sampNOHGD[c('Status', 'Prediction')], ci=T, of='thresholds')
roc.plot(rocNOHGD, 'No HGD')  
```

```{r, echo=F, fig.width=6, fig.height=6,eval=F}
ggplot(pg.sampNOHGD, aes(Prediction)) + geom_histogram(aes(fill=..x..), breaks=seq(0,1,0.1), show.legend = F) + scale_fill_distiller(palette = 'RdYlBu', name='P(P)') + labs(title='Prediction using model w/o HGD', y='n Samples', x='Probability') + plot.theme

@@

noHGD.pred.confidence = as.data.frame.matrix(table(pg.sampNOHGD$quants, pg.sampNOHGD$Status))
noHGD.pred.confidence = cbind(noHGD.pred.confidence, pg.sampNOHGD %>% group_by(quants) %>% dplyr::summarise ( 'mn'=mean(Prediction), 'sd'=sd(Prediction) ))
noHGD.pred.confidence$quant = rownames(noHGD.pred.confidence)

ft.fun<-function(NP,P) {
  print(paste(NP, P))
  f = fisher.test(rbind(cbind(NP,P),table(sum.patient.data$Status)))
  cbind.data.frame('p.value'=round(f$p.value, 5), 'odds.ratio'=round(f$estimate, 4))
}

noHGD.pred.confidence = noHGD.pred.confidence %>% as.data.frame.matrix %>% group_by(quant) %>% 
  dplyr::mutate( 'perc'=P/sum(NP,P),'p.value'=ft.fun(NP,P)$p.value, 'odds'=ft.fun(NP,P)$odds.ratio, 'conf'=ifelse(p.value < 0.05, '*', '') )

@@

ggplot(noHGD.pred.confidence, aes(mn, perc)) + 
  geom_rect(aes(xmin=r1, xmax=r2, ymin=0,ymax=1, fill=Risk), alpha=0.6) + scale_fill_manual(values=riskCols) +
  geom_point() + geom_text(aes(label=paste(r1,r2,sep='-')), nudge_x=0.05) +
  geom_errorbar(aes(ymin=ci.low, ymax=ci.high), width=0.01) + labs(x='Mean prediction', y='Ratio of `P` patients', title='Prediction calibration', subtitle='HGD excluded model') + plot.theme + theme(legend.position = 'none')

# Predict HGD samples
file = paste(cache.dir, 'nohgd.Rdata', sep='/')
if (file.exists(file)) {
  load(file, verbose=T)

@@

  hgd.pred = hgd.samples %>% group_by(Patient, Hospital.Research.ID, Status, Endoscopy.Year) %>% 
    dplyr::summarise( max=max(Pathology), pred=Prediction[which.max(Pathology)] )

@@

rm(plots, nzcoefs, fits)
```

@@

## All LOO

```{r echo=F, eval=F}

files = list.files('~/Data/Ellie/Analysis/5e6_arms_LOO', full.names = T)

@@

for (pt in sum.patient.data$Hospital.Research.ID) {
  ptF = grep(pt, files, value=T)
  load(grep('loo\\.Rdata', ptF, value=T), verbose=T)
  #load(grep('model_data\\.Rdata', ptF, value=T), verbose=T)
  performance[pt] = performance.at.1se
  pg.samp[[pt]] = info
}
rm(plots,coefs,nzcoefs,fits,info)

load("/Users/skillcoy/Data/Ellie/Analysis/5e6_arms_LOO/PR1_HIN_042_loo.Rdata", verbose=T)
load("/Users/skillcoy/Data/Ellie/Analysis/5e6_arms_LOO/PR1_HIN_042_model_data.Rdata", verbose=T)

load("/Users/skillcoy/Data/Ellie/Analysis/5e6_arms_LOO/AHM0320_loo.Rdata", verbose=T)
load("/Users/skillcoy/Data/Ellie/Analysis/5e6_arms_LOO/AHM0320_model_data.Rdata", verbose=T)

@@

```{r insilico_rocs, echo=F, message=F, warning=F, fig.height=8, fig.width=8}

load(file='~/Data/Ellie/Analysis/5e6_arms_all/insilico_splits.Rdata', verbose=T)

@@

insilico.rocs[['all']] = roc

@@

p = ggplot(df, aes(Specificity, Sensitivity,color=is)) + geom_line(size=1) + 
    geom_line(data=subset(df, is == 'all'), color='red', size=1) +
    scale_x_reverse() +
    scale_color_brewer(palette='Blues') +
    labs(title='ROC, all splits', x='Specificity (FPR)', y='Sensitivity (TPR)')  + plot.theme +
    theme(legend.position = 'none')
p
ggsave(filename='plots/splits_roc.png', plot=p, width=6, height=6, units='in', dpi=300)

@@

rr = ggplot(subset(df, is == 'all' | auc %in% c(range(auc))), aes(Specificity, Sensitivity)) + 
  geom_line(aes(color=factor(model)), size=1) + 
  geom_line(data=subset(df, auc == min(auc)), color='darkblue', linetype='dotted') + 
  geom_line(data=subset(df, auc == max(auc)), color='darkblue', linetype='dotted') +
  scale_x_reverse() + scale_color_brewer(palette='Paired') + 
  geom_label(data=subset(best, model == 'all' | auc %in% c(min(auc), max(auc)) ), aes(specificity, sensitivity, label=paste('AUC ',round(auc, 2), sep='\n'))) +
  labs(title='ROC', x='Specificity (1-FPR)', y='Sensitivity (TPR)')  + plot.theme +
  theme(legend.position = 'none')
rr  
ggsave(filename='plots/roc_CI.png', plot=rr, width=6, height=6, units='in', dpi=300)

@@

comp = ggplot(melt(subset(m, Measure == 'AUC'), measure.vars='value'), aes(x=model, y=value*100, ymin=CI.min*100, ymax=CI.max*100, group=Measure, fill=Measure)) + 
  geom_bar(stat='identity',position=dodge) + geom_errorbar(position=dodge, width=0.3, color='black' ) + 
  geom_text(aes(y=(value*100)-4, label=paste(round(value,2)*100,'%',sep='')), position=dodge, color='white')  +
  scale_fill_manual(values=cols, name='') + plot.theme + 
  theme(legend.position = 'none') + labs(x='', y='AUC', title='ROC Comparison') 
comp

ggsave('plots/auc/path_p53_cn_auc.png', plot=comp, width=2.7, height=3.7, units="in", limitsize=F, dpi = 300, scale=2)

@@

```{r back, echo=F, message=F, warning=F}
transform.by.time<-function(pt, info) {
  pt = pt[,c('Status','Patient','Hospital.Research.ID','Path.ID','PID','Pathology','Samplename','Final.Endoscopy','Endoscopy.Year','Prediction', 'OR')] %>% group_by(PID, Samplename) %>% dplyr::mutate(
    'months.before.final' = (Final.Endoscopy - Endoscopy.Year)*12)

  pt = arrange(merge(pt, subset(info, Hospital.Research.ID == pt$Hospital.Research.ID[1], select=c('Samplename','Block')), by='Samplename'), Pathology, PID, Endoscopy.Year)

  tb = with(pt, table(Endoscopy.Year, PID))
  tb[tb>0] = 1
  for (yr in names(which(rowSums(tb) > 1))) {
    tmp = pt[which(pt$Endoscopy.Year == yr),]
    tmp = tmp %>% group_by(PID) %>% summarise(max(Pathology))

    match = which(pt$Endoscopy.Year == yr & pt$PID == as.character(tmp[which.min(tmp$`max(Pathology)`), 'PID']))
    pt[match,'months.before.final'] = pt[match,'months.before.final'] + 6
  }

  pt$nl = (pt$Block-1)*2

  tb = table(pt[,c('months.before.final', 'PID')])
  for (i in 1:nrow(tb)) {
    pids = tb[i,]
    yrs = which(pids > 0)
    if (length(yrs) > 1) {
      for (j in 2:length(yrs)) 
        pt[which(pt$PID == names(yrs)[j]), 'nl'] = pt[which(pt$PID == names(yrs)[j]), 'nl'] + 1
    }
  }
  arrange(pt, -Endoscopy.Year, PID)
}

@@

pcal = ggplot(pred.confidence, aes(mn, perc)) + 
  geom_rect(aes(xmin=r1, xmax=r2, ymin=0,ymax=1, fill=Risk), alpha=0.6) + scale_fill_manual(values=riskCols, limits=levels(pred.confidence$Risk) ) +
  geom_point() + geom_abline(slope=1, intercept=0, color='grey39', linetype='dotted') + #geom_text(aes(label=paste(r1,r2,sep='-')), nudge_x=0.05) +
  geom_errorbar(aes(ymin=ci.low, ymax=ci.high), width=0.01) + labs(x='Mean prediction', y='Ratio of `P` patients', title='Prediction calibration') +
  scale_color_manual(values=riskCols,limits=levels(pred.confidence$Risk) ) + plot.theme + theme(legend.position = 'none')
pcal

ggsave(filename='plots/pred_calib.png', plot=pcal, width=2.2, height=2.2, units='in', dpi=300, scale=2)

@@

unadjRR = ggplot(back, aes(OR)) + geom_histogram(aes(fill=..x..), bins=15, show.legend = F) +
  #scale_fill_distiller(palette = 'RdYlBu', type='div',name='P(P)') +
  scale_fill_gradientn(colors = myPal,  name='') + 
  labs(y='n Samples', x='Relative Risk', title='Unadjusted relative risk') + plot.theme
unadjRR
ggsave(filename='plots/unadj_relrisk.png', plot=unadjRR, width=5.5, height=4.3, units='in', dpi=300)

@@

ggsave(filename='plots/feature_jacknife.png', plot=p, width=6, height=8, unit='in', dpi=300)

@@

rp = ggplot(riskPath, aes(Pathology, ratio*100)) + 
  geom_bar(aes(group=Risk, fill=Risk),stat='identity', position=position_stack()) + 
  scale_fill_manual(values=riskCols, limits=levels(riskPath$Risk)) + 
  scale_color_manual(values=c('white','black','white'), limits=levels(pred.confidence$Risk) ) +
  geom_text(data=pathTotal, aes(label=paste('n=',nPath,sep=''), x=Pathology, y=101.5), position=position_stack(), size=4) +
  geom_text(data=subset(riskPath, ratio>0), aes(group=Risk,label=paste(ratio*100,'%',sep=''), color=Risk), position=position_stack(vjust = 0.5), size=5, show.legend = F) +
  facet_grid(~Status, scales='free_x', space = 'free_x') + plot.theme + 
  labs(title='Samples predicted by pathology', y='% Predicted', x='Pathology') + theme(legend.position = 'bottom')
rp

ggsave(filename='plots/pred_by_path.png', plot=rp, width=5.8, height=4.5, units='in', dpi=300, scale=1.5)

@@

p53Path$p53.Status = factor(p53Path$p53.Status, ordered = T)
p53Path = arrange(p53Path, Pathology, p53.Status)
col = brewer.pal(11, "PRGn")[c(10,2)]
rp = ggplot(p53Path, aes(Pathology, ratio*100, group=Pathology, fill=p53.Status)) + geom_bar(stat='identity', position = position_stack()) +
  scale_fill_manual(values=col, name='p53 IHC', labels=c('Normal','Aberrant')) + 
  geom_text(aes(label=paste('n=',nPath,sep=''), x=Pathology, y=102)) +
  geom_text(aes(label=paste(ratio*100,'%',sep='')), position=position_stack(vjust = 0.5), size=5,color='white') +
  plot.theme + labs(title='Samples predicted by aberrant p53 IHC', y='% Predicted', x='Pathology') + theme(legend.position = 'bottom')
rp
ggsave(filename='plots/p53_pred.png', plot=rp, width=4.5, height=3.6, units='in', dpi=300, scale=1.5)

@@

endos = back %>% group_by(Status, months.before.final, brks, PID) %>% dplyr::summarise(
  'pred.endo'=ifelse(length(which(Risk >= 'Moderate')) > 0, 1, 0), 
  'n.endos'=length(unique(PID)), 
  'n.samples'=length(Samplename),
  'mean.pred'=mean(Prediction),
  'pred.samples'=length(which(Risk >= 'Moderate')),
  'max.path'=max(Pathology),
  'Moderate'=length(which(Risk == 'Moderate')),
  'High'=length(which(Risk == 'High'))
  )

@@

endos.m = subset(endos, Status == 'P') %>% group_by(brks) %>% dplyr::summarise( 
  'pred.r'=sum(pred.endo)/sum(n.endos),'n.samp'=sum(n.samples), 'n.endo'=sum(n.endos), 
  'mod.r'=sum(Moderate)/sum(n.samples), 'high.r'=sum(High)/sum(n.samples))

@@

endos.m = arrange(merge(endos.m,tmp), brks)

rp = ggplot(endos.m, aes(x=brks, y=pred.r*100, ymin=(pred.r-`sem(P)`)*100, ymax=(pred.r+`sem(P)`)*100, group=brks)) + ylim(0,105) +
  geom_bar(stat='identity', fill='orangered4', show.legend = F) + 
  geom_errorbar(color='grey21', width=0.2) +
  geom_text(aes(y=3, label=paste(round(pred.r*100), '%', sep='')), color='white') +
  geom_text(aes(y=pred.r*100-5, label=paste('n=',n.endo,'\ns=',n.samp,sep='')), color='white') + 
  labs(x='Months prior to endpoint', y='Percent predicted', title='Predictions prior to final endoscopy', subtitle='n = #endoscopies, s = #samples') +
  plot.theme + theme( axis.text.x=element_text(angle=45,hjust=1), text=element_text(size=12)) 
rp
ggsave(filename='plots/pred_endo.png', plot=rp, width=5.6, height=4.4, units='in', dpi=300, scale=1.5)

@@

```{r, echo=F, warning=F, message=F, fig.width=5, fig.height=5 }
back$nl = factor(back$nl, ordered=T)
back$Risk = factor(back$Risk, levels=c('Low','Moderate','High'), ordered = T)
back$Patient = factor(as.numeric(back$Patient), ordered = T)

@@

p = pred.tiles(subset(back, Patient %in% c(6)), ncol=1, cols=myPal, probs=T) + labs(title='')
tilesLegend = get.legend(p)
p + theme(legend.position = 'none')

@@

```{r, echo=F, message=F, warning=F, width=26, height=8}
back = back %>% group_by(Patient) %>% dplyr::mutate( 'pred.ratio'=round(sum(Prediction)/length(Prediction),2))

>>>>

all.patient.info = read.patient.info(patient.file, demo.file, set='all')$info
all.patient.info = subset(all.patient.info, Hospital.Research.ID != 'AHM0320')

@@

@@

ggsave('plots/per_sample_cx.png', plot=pv, width=4, height=4, units='in', dpi=300)
ggsave('plots/per_sample_cx_roc.png', plot=roc.plot(rocX, title='') + plot.theme , width=4, height=4, units='in', dpi=300)

@@

```{r model, message=F, warning=F, echo=F, fig.height=16, fig.width=16}
#folds = 10; splits = 5 
# file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
# if (file.exists(file)) {
#   message(paste("loading", file))
#   load(file, verbose=T)
# } else {
#   stop(paste("No model file", file, "found"))
# }

@@

```{r model_perf, message=F, warning=F, echo=F, fig.width=6, fig.height=6}
folds = 10; splits = 5 
classification.rate = as_tibble(matrix(ncol=5, nrow=0, dimnames=list(c(), c('model','mean','sme','stable.coef', 'all.coef'))))

@@

ggsave('plots/performance/ridge_v_lasso.png', plot=mp$plot, width=3.5, height=3.5, units='in', dpi=300, scale=1.2)

@@

ggsave('plots/performance/ridge_v_lasso_nohgd.png', plot=mp$plot, width=3.5, height=3.5, units='in', dpi=300, scale=1.2)

@@

@@

```{r loonohgd, echo=F, message=F, warning=T, fig.width=5, fig.height=5}
file = paste(cache.dir, 'loonohgd.Rdata', sep='/')
if (file.exists(file)) {
  message(paste("loading file", file))
  load(file, verbose=T)
} else {
  stop(paste("No file", file, "found"))
}

pg.sampNOHGD = do.call(rbind, pg.sampNOHGD)
perf = cbind.data.frame(perf=performance.at.1se, pt=unique(pg.sampNOHGD$Patient))

p = ggplot(melt(perf), aes(x=variable,y=value, group=variable, fill=variable)) +
    geom_boxplot(fill='lightcoral', color='darkgrey', outlier.fill=NA, outlier.color=NA, size=0.8) +
    geom_jitter(color='grey39', width=0.3) + plot.theme + 
    geom_label(data=melt(subset(perf, perf %in% range(perf) )), aes(x=variable, y=value, label=round(value, 3)), fontface='bold', fill='lightcoral') +
    theme(legend.position = 'none') + labs(x='LOO model', y='Classification rate', title='Performance for LOO', subtitle='HGD samples excluded model')
p

ggsave('plots/performance/loo_perf_nohgd.png', plot=p, width=3, height=3, units='in', dpi=300)

rocNOHGD = pROC::roc(Status ~ Prediction, data=pg.sampNOHGD[c('Status', 'Prediction')], ci=T, of='thresholds')

```{r, echo=F, fig.width=6, fig.height=6}
hgdloo = ggplot(pg.sampNOHGD, aes(Prediction)) + geom_histogram(aes(fill=..x..), breaks=seq(0,1,0.1), show.legend = F) + scale_fill_distiller(palette = 'RdYlBu', name='P(P)') + labs(title='Prediction using model w/o HGD', y='n Samples', x='Probability') + plot.theme
hgdloo

ggsave('plots/prediction_nohgd.png', plot=hgdloo, width=5, height=5, units='in', dpi=300)

@@

noHGD.pred.confidence = as.data.frame.matrix(table(pg.sampNOHGD$quants, pg.sampNOHGD$Status))
noHGD.pred.confidence = cbind(noHGD.pred.confidence, pg.sampNOHGD %>% group_by(quants) %>% summarise ( 'mn'=mean(Prediction), 'sd'=sd(Prediction) ))
noHGD.pred.confidence$quant = rownames(noHGD.pred.confidence)

ft.fun<-function(NP,P) {
  f = fisher.test(rbind(cbind(NP,P),table(sum.patient.data$Status)))
  cbind.data.frame('p.value'=f$p.value, 'odds.ratio'=f$estimate)
}

noHGD.pred.confidence = noHGD.pred.confidence %>% as.data.frame.matrix %>% group_by(quant) %>% mutate( 'perc'=P/sum(NP,P),'p.value'=ft.fun(NP,P)$p.value, 'odds'=ft.fun(NP,P)$odds.ratio, 'conf'=ifelse(p.value < 0.05, '*', '') )

@@

nohgdcalib = ggplot(noHGD.pred.confidence, aes(mn, perc)) + 
  geom_rect(aes(xmin=r1, xmax=r2, ymin=0,ymax=1, fill=Risk), alpha=0.6) + scale_fill_manual(values=riskCols) +
  scale_x_continuous(limits=c(0,1), breaks = cuts, labels = c(0.0, '','','','','',0.6,'',0.8,'',1.0)) +
  geom_point() + 
  geom_errorbar(aes(ymin=ci.low, ymax=ci.high), width=0.01) + labs(x='Mean prediction', y='Ratio of `P` patients', title='Prediction calibration', subtitle='HGD excluded model') + plot.theme + theme(legend.position = 'none')
nohgdcalib

ggsave('plots/calib_nohgd.png', plot=nohgdcalib, width=5, height=5, units='in', dpi=300)
# Predict HGD samples
file = paste(cache.dir, 'nohgd.Rdata', sep='/')
if (file.exists(file)) {
  load(file, verbose=T)

@@

  hgd.pred = hgd.samples %>% group_by(Patient, Hospital.Research.ID, Status, Endoscopy.Year) %>% 
    summarise( max=max(Pathology), pred=Prediction[which.max(Pathology)] )

@@

rm(plots, performance.at.1se,coefs, nzcoefs, fits)
```

ggsave('plots/performance/no_lgd_perf.png', plot=mp$plot, height=3.5, width=3.5, units='in', dpi=300, scale=1.2)

@@

```{r echo=F, eval=F}
## All LOO
files = list.files('~/Data/Ellie/Analysis/5e6_arms_LOO', full.names = T)

@@

for (pt in sum.patient.data$Hospital.Research.ID) {
  ptF = grep(pt, files, value=F)
  load(grep('loo\\.Rdata', ptF, value=T), verbose=T)
  #load(grep('model_data\\.Rdata', ptF, value=T), verbose=T)
  performance[pt] = performance.at.1se
  pg.samp[[pt]] = info
}
rm(plots,coefs,nzcoefs,fits,info)

#load("/Users/skillcoy/Data/Ellie/Analysis/5e6_arms_LOO/PR1_HIN_042_loo.Rdata", verbose=T)
#load("/Users/skillcoy/Data/Ellie/Analysis/5e6_arms_LOO/PR1_HIN_042_model_data.Rdata", verbose=T)

#load("/Users/skillcoy/Data/Ellie/Analysis/5e6_arms_LOO/AHM0320_loo.Rdata", verbose=T)
#load("/Users/skillcoy/Data/Ellie/Analysis/5e6_arms_LOO/AHM0320_model_data.Rdata", verbose=T)

@@

@@

load(file=paste(cache.dir, 'insilico_splits.Rdata', sep='/'), verbose=T)
insilico.rocs[['all']] = roc

@@

p = ggplot(df, aes(Specificity, Sensitivity,color=is)) + geom_line(size=1) + 
    geom_line(data=subset(df, is == 'all'), color='red', size=1) +
    scale_x_reverse() +
    scale_color_brewer(palette='Blues') +
    labs(title='ROC, all splits', x='Specificity (FPR)', y='Sensitivity (TPR)')  + plot.theme +
    theme(legend.position = 'none')
p
ggsave(filename='plots/auc/splits_roc.png', plot=p, width=6, height=6, units='in', dpi=300)

@@

rr = ggplot(subset(df, is == 'all' | auc %in% c(range(auc))), aes(Specificity, Sensitivity)) + 
  geom_line(aes(color=factor(model)), size=1) + 
  geom_line(data=subset(df, auc == min(auc)), color='darkblue', linetype='dotted') + 
  geom_line(data=subset(df, auc == max(auc)), color='darkblue', linetype='dotted') +
  scale_x_reverse() + scale_color_brewer(palette='Paired') + 
  geom_label(data=subset(best, model == 'all' | auc %in% c(min(auc), max(auc)) ), aes(specificity, sensitivity, label=paste('AUC ',round(auc, 2), sep='\n'))) +
  labs(title='ROC CI', x='Specificity (1-FPR)', y='Sensitivity (TPR)')  + plot.theme +
  theme(legend.position = 'none')
rr  
ggsave(filename='plots/auc/roc_CI.png', plot=rr, width=4, height=4, units='in', dpi=300, scale=1.5)

@@

comp = ggplot(melt(subset(m, Measure == 'AUC'), measure.vars='value'), aes(x=model, y=value*100, ymin=CI.min*100, ymax=CI.max*100, group=Measure, fill=Measure)) + 
  ylim(0,100) + 
  geom_bar(stat='identity',position=dodge) + geom_errorbar(position=dodge, width=0.3, color='black' ) + 
  geom_text(aes(y=(value*100)-4, label=paste(round(value,2)*100,'%',sep='')), position=dodge, color='white')  +
  scale_fill_manual(values=cols, name='') + plot.theme + 
  theme(legend.position = 'none', axis.text.x = element_text(hjust=1,angle=45)) + labs(x='', y='AUC', title='') 
comp

ggsave('plots/auc/path_p53_cn_auc.png', plot=comp, width=3.1, height=3.9, units="in", limitsize=F, dpi = 300, scale=1)

@@

```{r back, echo=F, message=F, warning=F}
transform.by.time<-function(pt, info) {
  pt = pt[,c('Status','Patient','Hospital.Research.ID','Path.ID','PID','Pathology','Samplename','Final.Endoscopy','Endoscopy.Year','Prediction', 'OR')] %>% group_by(PID, Samplename) %>% dplyr::mutate('months.before.final' = (Final.Endoscopy - Endoscopy.Year)*12)
  pt = dplyr::arrange(merge(pt, subset(info, Hospital.Research.ID == pt$Hospital.Research.ID[1], select=c('Samplename','Block')), by='Samplename'), Pathology, PID, Endoscopy.Year)

  tb = with(pt, table(months.before.final, PID))
  tb[tb>0] = 1
  for (mnths in names(which(rowSums(tb) > 1))) {
    tmp = pt[which(pt$months.before.final == mnths),]
    tmp = arrange(tmp, rev(Pathology), PID)

    mt = tmp$months.before.final[1] 
    interval = length(tmp$Endoscopy.Year)/(length(unique(tmp$PID))-1)
    for (pid in unique(tmp$PID)) {
      rows = which(tmp$PID == pid)
      if (length(grep('HGD|IMC',tmp[rows,'Pathology'])) > 0 & mt != 0) {
        mt = 0
        tmp$months.before.final[rows] = mt
      } else {
        mt = mt + interval
        tmp$months.before.final[rows] = mt
      }
    }
    pt[which(pt$months.before.final == mnths),] = tmp
  }

  pt$nl = (pt$Block-1)*2
  arrange(pt, -Endoscopy.Year, PID)
}

@@

pcal = ggplot(pred.confidence, aes(mn, perc)) + 
  geom_rect(aes(xmin=r1, xmax=r2, ymin=0,ymax=1, fill=Risk), alpha=0.6) + scale_fill_manual(values=riskCols, limits=levels(pred.confidence$Risk) ) +
  geom_point() + geom_abline(slope=1, intercept=0, color='grey39', linetype='dotted', size=0.5) + #geom_text(aes(label=paste(r1,r2,sep='-')), nudge_x=0.05) +
  geom_errorbar(aes(ymin=ci.low, ymax=ci.high), size=0.5, width=0.01) + labs(x='P(P)', y='NP:P', title='') +
  scale_color_manual(values=riskCols,limits=levels(pred.confidence$Risk) ) + 
  scale_x_continuous(limits=c(0,1), breaks = cuts, labels = c(0.0, '','',0.3,'','','',0.7,'','',1.0)) +
  plot.theme + theme(legend.position = 'none')
pcal

ggsave(filename='plots/pred_calib.png', plot=pcal, width=2.2, height=2.2, units='in', dpi=300, scale=1.5)

@@

unadjRR = ggplot(back, aes(OR)) + geom_histogram(aes(fill=..x..), bins=15, show.legend = F) +
  scale_fill_gradientn(colors = myPal,  name='') + 
  labs(y='n Samples', x='Relative Risk', title='Unadjusted relative risk') + plot.theme
unadjRR
ggsave(filename='plots/unadj_relrisk.png', plot=unadjRR + labs(title=''), width=5.5, height=4.3, units='in', dpi=300)

@@

ggsave(filename='plots/feature_jacknife.png', plot=p, width=6, height=10, unit='in', dpi=300)

@@

rp = ggplot(riskPath, aes(Pathology, ratio*100)) + 
  geom_bar(aes(group=Risk, fill=Risk),stat='identity', position=position_stack()) + 
  scale_fill_manual(values=riskCols, limits=levels(riskPath$Risk)) + 
  scale_color_manual(values=c('white','black','white'), limits=levels(pred.confidence$Risk) ) +
  geom_text(data=pathTotal, aes(label=paste('n=',nPath,sep=''), x=Pathology, y=102), position=position_stack()) +
  geom_text(data=subset(riskPath, ratio>0), aes(group=Risk,label=paste(ratio*100,'%',sep=''), color=Risk), position=position_stack(vjust = 0.5), show.legend = F) +
  facet_grid(~Status, scales='free_x', space = 'free_x') + plot.theme + 
  labs(title='Samples predicted by pathology', y='% Predicted', x='Pathology') + theme(legend.position = 'none')
rp

ggsave(filename='plots/pred_by_path.png', plot=rp, width=5.8, height=4.5, units='in', dpi=300, scale=1)

@@

p53Path$p53.Status = factor(p53Path$p53.Status, ordered = T)
p53Path = arrange(p53Path, Pathology, p53.Status)
col = brewer.pal(11, "PRGn")[c(10,2)]
rp = ggplot(p53Path, aes(Pathology, ratio*100, group=Pathology, fill=p53.Status)) + geom_bar(stat='identity', position = position_stack()) +
  scale_fill_manual(values=col, name='p53 IHC', labels=c('Normal','Aberrant')) + 
  geom_text(aes(label=paste('n=',nPath,sep=''), x=Pathology, y=103)) +
  geom_text(aes(label=paste(ratio*100,'%',sep='')), position=position_stack(vjust = 0.5), size=5,color='white') +
  plot.theme + labs(title='', y='% Predicted', x='Pathology') + theme(legend.position = 'bottom')
rp
ggsave(filename='plots/p53_pred.png', plot=rp, width=4.5, height=3.6, units='in', dpi=300, scale=1.2)

@@

endos = back %>% group_by(Status, months.before.final, brks, PID) %>% dplyr::summarise(
  'pred.endo'=ifelse(length(which(Risk > 'Moderate')) > 0, 1, 0), 
  'n.endos'=length(unique(PID)), 
  'n.samples'=length(Samplename),
  'mean.pred'=mean(Prediction),
  'pred.samples'=length(which(Risk > 'Moderate')),
  'max.path'=max(Pathology),
  'Moderate'=length(which(Risk == 'Moderate')),
  'High'=length(which(Risk == 'High'))
  )

endos.mod = back %>% group_by(Status, months.before.final, brks, PID) %>% dplyr::summarise(
  'pred.endo'=ifelse(length(which(Risk >= 'Moderate')) > 0, 1, 0), 
  'n.endos'=length(unique(PID)), 
  'n.samples'=length(Samplename),
  'mean.pred'=mean(Prediction),
  'pred.samples'=length(which(Risk >= 'Moderate')),
  'max.path'=max(Pathology),
  'Moderate'=length(which(Risk == 'Moderate')),
  'High'=length(which(Risk == 'High'))
  )

@@

endos.mod$brks = cut(endos.mod$months.before.final, include.lowest = T, ordered_result = T,breaks=sq,labels=mlab)
endos.mod$brks = with(endos.mod, factor(brks, rev(levels(brks)))) # invert order

endos.r = subset(endos, Status == 'P') %>% group_by(brks) %>% dplyr::summarise( 
  'pred.r'=sum(pred.endo)/sum(n.endos),'n.samp'=sum(n.samples), 'n.endo'=sum(n.endos), 
  'mod.r'=sum(Moderate)/sum(n.samples), 'high.r'=sum(High)/sum(n.samples))

endos.m = subset(endos.mod, Status == 'P') %>% group_by(brks) %>% dplyr::summarise( 
  'pred.r'=sum(pred.endo)/sum(n.endos),'n.samp'=sum(n.samples), 'n.endo'=sum(n.endos), 
  'mod.r'=sum(Moderate)/sum(n.samples), 'high.r'=sum(High)/sum(n.samples))

@@

endos.r = arrange(merge(endos.r,tmp), brks)
endos.m = arrange(merge(endos.m,tmp), brks)

rp = ggplot(endos.r, aes(x=brks, y=pred.r*100, ymin=(pred.r-`sem(P)`)*100, ymax=(pred.r+`sem(P)`)*100, group=brks)) + ylim(0,105) +
#  geom_bar(data=endos.m, aes(x=brks, y=pred.r*100, group=brks), stat='identity', fill=riskCols[2], show.legend = F, alpha=0.5) +
#  geom_text(data=endos.m, aes(y=round(pred.r*100+2), label=paste(round(pred.r*100), '%', sep='')), color='grey45') +
  geom_bar(stat='identity', fill=riskCols[3], show.legend = F) + 
  geom_errorbar(color='grey21', width=0.2) +
  geom_text(aes(y=3, label=paste(round(pred.r*100), '%', sep='')), color='white') +
  geom_text(aes(y=pred.r*100-8, label=paste('n=',n.endo,'\ns=',n.samp,sep='')), color='white') + 
  labs(x='Months prior to endpoint', y='Percent predicted', title='Predictions prior to final endoscopy', subtitle='n = #endoscopies, s = #samples') +
  plot.theme + theme( axis.text.x=element_text(angle=45,hjust=1) ) 
rp
ggsave(filename='plots/pred_endo.png', plot=rp, width=5.8, height=4.5, units='in', dpi=300, scale=1)

@@

```{r heatmaps, echo=F, warning=F, message=F, fig.width=5, fig.height=5 }
back$nl = factor(back$nl, ordered=T)
back$Risk = factor(back$Risk, levels=c('Low','Moderate','High'), ordered = T)
back$Patient = factor(as.numeric(back$Patient), ordered = T)

@@

@@

```{r, echo=F, message=F, warning=F, width=26, height=8}
back = back %>% group_by(Patient) %>% dplyr::mutate( 'pred.ratio'=round(sum(Prediction)/length(Prediction),2), 'n.endos'=length(unique(months.before.final)), 'max.block'=max(Block))
back = transform(back, Patient=reorder(Patient, max.block))
back$Block = factor(back$Block, ordered = T)

pred.tiles(subset(back, Patient %in% c(1,20,34,21,3, 19)), cols=myPal, probs=T, space='free', scales='free_x', drop=T) + theme(legend.position = 'none')
pred.tiles(subset(back, Patient %in% c(1,20,34,21,3, 19)), cols=riskCols, probs=F, space='free', scales='free_x', drop=T) + theme(legend.position = 'none')

w = 11; h = 1.2; s = 1
